Search for Spatially Extended \fermi-LAT Sources Using Two Years of Flight Data

### 2011-09-22 extended_catalog/paper/extended_catalog.tex
improving draft

% Changes from v2.0 to v3.0

% * Move section 'Dual Localization' to its own
%   section after 'Extended Source Detection Threshold'
%   and rename it 'Testing Against Source Confusion'.
%   The section is significantly expanded with a new
%   monte carlo study fitting extended sources as
%   two point-like sources, and vice versa
%
% * Major rewrite of ``Extended Source Search Method'' section.
%   This section now more correctly explains how the
%   analysis search was done as well as cautions the
%   reader about what caveats may apply to the analysis.

% * Major rewrite of the sections on the new extended sources
%   to better describe why various point sources were removed
%   and why some were kept using multiwavelenth information.

% * Move AGN validation section earlier in the text
%   so that it is the first section analyzing actual LAT 
% * Move section on reanalyzing 2FGL sources
%   to earlier in the paper to be right after the AGN
%   validation.
% * Move earlier section discussing systematic errors
% * better description of how two of the 'new extended sources'
%   already have dedicated publications

% * Remake 2FGL J1632.4-4753c smoothed counts map plot to show
%   before + after background subtraction
%
% * Reordered Gamma-Cygni section to first discuss the pulsar.
%
% * For HESS J1837, smooth image by a 0.15 degree Gaussain
%   instead of a 0.1 degree gausian
%   kernel to surpress some of the noise which is likely spurious.
%   Also, add in some lower level contour from HESS data
%   to better show how wide the HESS emission is.
%
% * Rewrote the 'preview of what's to come section
%   in the introduction to have the right order.
%
% * Cleaned up the Comparing source size section.
% * Better description of converting TSext into sigma.
% * better warning about the limits of the alternate diffuse model.
% * Fixed figure order.
% * made tickmarks white in plots.
% * replace all occurrences of ts_inc with ts_2pts
% * changed zoom of IC443 res tsmap to be easier to view.
% * Added LAT collaboration citation of W30.
%
% * Added acknowlegements section to end of the paper
% * Move HESS SEDs to earlier in paper to fix order.
% * Fix some occurances of chi^2 to correctly be chi^2_1
% * Better description of how the alternate diffuse model
%   is different from galprop
% * Add galprop citation
% * Add AGILE citation Gamma Cygni Pulsar citation.

@@

\newif\ifcolorfigure
\colorfiguretrue
%\colorfigurefalse

@@

\author{
J.~Lande\altaffilmark{2},
S.~Funk\altaffilmark{2},
M.~Ackermann (???),
%A.~Allafort\altaffilmark{2},
%M.-H.~Grondin\altaffilmark{53},
% Keith Bechtol
% Marshall Roth?
% Marianne Lemoine-Goumard
% Francesco Giordano
% Toby Burnett
% Jean Ballet
% Matthew Kerr
\ldots
\altaffiltext{2}{W. W. Hansen Experimental Physics Laboratory, Kavli Institute for Particle Astrophysics and Cosmology, Department of Physics and SLAC National Accelerator Laboratory, Stanford University, Stanford, CA 94305, USA}
\altaffiltext{53}{Institut f\"ur Astronomie und Astrophysik, Universit\"at T\"ubingen, D 72076 T\"ubingen, Germany}
}

\begin{abstract}
Spatial extension is an important characteristic for correctly
associating LAT sources with their counterparts at other wavelengths
and for obtaining an unbiased model of their spectra.  We present a new
method for quantifying the spatial extension of sources with the Large
Area Telescope (LAT), the primary science instrument on the {\em \fermi
Gamma-ray Space Telescope} (\fermi).  We perform a series of Monte Carlo
simulations to validate this tool and calculate the LAT threshold for
detecting the spatial extension of sources.  We then test all sources
in the second \fermi-LAT catalog (2FGL) for extension. 
\newtext{
We report the
detection of seven spatially extended sources in addition to the twelve
spatially extended sources reported in 2FGL and two other sources studied
in dedicated publications.
}
\end{abstract}

@@

\newtext{
Previous analysis of extended LAT sources were performed as dedicated
studies of individual sources, so we expect that a systematic scan of
all LAT detected sources could uncover additional spatially extended
sources.  For these reasons, in Section~\ref{analysis_methods_section}
we present a new systematic method for analyzing spatially extended
LAT sources.  In Section~\ref{monte_carlo_validation}, we demonstrate
that this method can be used to test the statistical significance of the
extension of a LAT source and in Section~\ref{extension_sensitivity}
we calculate the LAT detection threshold to resolve the extension
of a source.  In Section~\ref{dual_localization_method}, we
study the ability of the LAT to distinguish between spatially
extended sources and confusion of nearby point-like sources.
In Section~\ref{test_2lac_sources} we further demonstrate that our
detection method does not misidentify point-like sources as being
extended by testing the extension of active galactic nuclei (AGN)
believed to be unresolvable.  In Section~\ref{validate_known},
we systematically reanalyze the twelve extended sources included
in 2FGL and in Section~\ref{systematic_errors_on_extension}
we describe a way to estimate systematic errors on the extension of a source.
In Section~\ref{extended_source_search_method}, we
describe a search for new spatially extended LAT sources. Finally,
in Section~\ref{new_ext_srcs_section} we present the detection of the
extension of nine spatially extended sources that were reported in 2FGL
but as being point-like.  Two of these have been previously analyzed in
dedicated publications.
}

@@

\newtext{
We considered two different models to use for the
surface brightness profile for extended sources, either a 2D Gaussian
profile
\begin{equation}\label{gauss_pdf}
  I_\text{Gaussian}(x,y)=\tfrac{1}{2\pi\sigma^2}\exp\left(-(x^2+y^2)/2\sigma^2\right)
\end{equation}
or a uniform surface brightness profile
\begin{equation}\label{disk_pdf}
  I_\text{disk}(x,y)=
  \begin{cases}
    \frac{1}{\pi\sigma^2} & x^2+y^2\le\sigma^2 \\
    0                      & x^2+y^2>\sigma^2.
  \end{cases}
\end{equation}
Although these shapes are significantly different, 
Figure~\ref{compare_disk_gauss} shows the their PDF 
are similar for a source that has a 
0\fdg5
radius
typical of LAT detected extended sources.
To
allow a valid comparison between the Gaussian and the uniform surface
brightness morphologies,
we defined here the source size (0\fdg5) 
as the radius containing
68\% of the intensity ($\rsixeight$).  $\rsixeight_\text{,Gaussian}=1.51\sigma$
and $\rsixeight_\text{,disk}=0.82\sigma$ where $\sigma$
is defined in Equation~\ref{gauss_pdf} and
Equation~\ref{disk_pdf} respectively.
For this example, where $\rsixeight=0\fdg5$, 
$\sigma_\text{disk}=0.61\degree$ and 
$\sigma_\text{Gaussian}=0.33\degree$.

For smaller and fainter extended sources, the differences in the PDF
for different spatial models are lost in the noise and the LAT is not
sensitive to the detailed spatial structure of the sources.  Therefore,
in our search for extended sources we use only a radially-symmetric
uniform surface brightness as our spatial extension shape. We quote the
radius of the edge ($\sigma$) as the size of the source.
}

@@

Figure~\ref{index_sensitivity} shows the threshold for sources of four
spectral indices from 1.5 to 3 and extensions varying from $\sigma=0\fdg1$
to $2\fdg0$.  
The LAT flux threshold for a significant detection
of source extension drops quickly with
increasing source size and reaches a minimum around 0\fdg5. 
\newtext{
The threshold is high for small extensions because the
source becomes
small compared to the PSF. The threshold increases
for large extended sources because the source becomes
increasingly diluted by the background.
} Figure~\ref{index_sensitivity} shows
the threshold using photons with energies between 100 \mev and 100 \gev
and also using only photons with energies between 1 \gev and 100 \gev.
Except for very large or very soft
sources, our detection threshold is
not substantially improved by including photons with energies between 100 \mev and
1 \gev.  This is also demonstrated in Figure~\ref{four_plots_ic443}
which shows \tsext for the SNR IC443 computed independently in twelve
energy bins between 100 \mev and 100 \gev. For IC443, which has a
spectral index $\sim2.4$ and an extension $\sim0\fdg35$, 
almost the entire increase in likelihood
modeling the source as being extended comes
from energies above 1 \gev.  Furthermore, other systematic errors
become increasingly important at low energy. For our extension search
(Section~\ref{extended_source_search_method}),
we therefore used only photons with energies above 1 \gev.

@@

We assessed the power of the $\tsext>\tsinc$ test using a Monte Carlo
study.  We simulated one spatially extended sources and fit it as both
an extended source and two point-like sources.  We then simulated two
point-like sources and fit them with the same two hypothesis. By comparing
the distribution of \tsinc and \tsext for the two cases, we evaluated
how well comparing \tsext to \tsinc is at rejecting the possibility of
source confusion as well as how likely it is to incorrectly reject an
extended source as being extended.  All sources were simulated using
the same two-year time range included in 2FGL and all fit values were
calculated using \pointlike. The simulations were performed against
an background 10 times the isotropic background measured by EGRET,
representative of the background near the Galactic plane.

We did this study first in the energy range from 1 \gev to 100 \gev by
simulating extended sources of flux $4\times10^{-9} \phflux$ integrated
from 1 \gev to 100 \gev and a power-law spectral model with
spectral index 2.  This flux was picked to be representative of the
new extended sourced that were discovered in the following analysis
when looking in the 1 \gev and 100 \gev energy range
(see Section~\ref{new_ext_srcs_section}).
We simulated these sources using a uniform surface brightness spatial
model with extensions varying
up to $1\degree$.  Figure~\ref{confusion_plot}a shows 
the distribution of \tsinc and \tsext as a
function of the simulated extension of the source
for 200 statistically independent simulations.

Figure~\ref{confusion_plot}c shows the same analysis but when
fitting simulations of two point-like sources each with half of the
flux of the spatially extended source but the same spectral index.
Finally, figure~\ref{confusion_plot}e shows the same analysis with each
point-like source having the same flux but different spectral indices.
One point-like source had a spectral index of 1.5 and the other an index
of 2.5.

@@

The distributions show that for simulated
extended sources with typical
sizes ($\sigma\sim0\fdg5$), one can often obtain almost as good an
increase in likelihood
when fitting the extended source as two point-like sources
($\tsext\sim\tsinc$).
This is because although
two point-like sources represent an incorrect spatial model, 
the two point-like sources have four additional degrees
of freedom (two spatial and two spectral parameters)
than the one extended source (one extension parameter)
and can therefore easy model
statistical fluctuations.
This effect is most pronounced when using energies from 1
\gev to 100 \gev where the PSF is even larger.  
On the other hand,
when testing the two simulated point-like sources for extension our test
did a better job at separating the distribution of \tsinc and \tsext
and found that \tsinc was typically much larger than \tsext.
This is because the one additional extension degree of freedom does
now as easily fit fit statistical fluctuations.

From this, we conclude that the $\tsext>\tsinc$ test is reasonably
powerful at discarding cases in which the true emission is derived
from two point-like sources instead of from a spatially extended
source. On the other hand, because of the large values of \tsinc
in figure~\ref{confusion_plot}a and figure~\ref{confusion_plot}a when
fitting spatially extended sources, we caution that even in the case of
their being spatially extended emission the fit of two point-like sources
can often seem (by comparing the likelihoods) to be almost as good.

Finally, figure~\ref{index_difference} shows the absolute value of the
difference of the fit spectral indices for the case of fitting simulated
spatially extended sources as two point-like sources. The plot compares
the fit in the 1 \gev to 100 \gev energy range and in the 10 \gev to
100 \gev energy range. An error on the difference in spectral index is
calculated by adding in quadrature the statistical error on the spectral
index.  This plot shows that there can be a large difference in the
spectral index of the two fit point-like sources which is often greater
than the statistical errors on the difference in spectral index. This is
because the fit values the errors on an incorrect model are not expected
to be meaningful. Therefore, the difference in spectral index when fitting
an extended source candidate as two point-like sources can not be easily
used to help determine whether a source is actually spatially extended
instead of the emission from two confused point-like sources.

From this Monte Carlo study, we can see the limits of a LAT analysis of
spatially extended sources.  Section~\ref{monte_carlo_validation} showed
that we have a statistical test which can rigorously find sources for
which the spatial emission is not well described by the PSF.  But for
the extended sources we are searching for, this extension test does
not uniquely prove that the emission originates from spatially extended
emission instead of from multiple unresolved sources.  Demanding that
$\tsext>\tsinc$ is a reasonably powerful test to avoid cases of simple
confusion with two point-like sources. But it could always be the
case that an extended source candidate is actually the superposition
of multiple point-like or extended sources which could be resolved
with deeper observations of the region.  Of course, there is nothing
about this conclusion unique to LAT analysis.  Its just that the 
larger PSF of the LAT and the density of sources expected to be \gev emitters in
the galactic plane makes this issue more significant for LAT analysis.
When possible, multiwavelength information should be used to help decide
what is qualitatively the best sky model.

@@

\newtext{
2FGL\,J0851.7$-$4635 was found by our search to be an extended source
candidate in the 10 \gev to 100 \gev energy range and is spatially
coincident with the SNR Vela Jr.  Figure~\ref{Vela_Jr} shows a counts
map of the source.  Overlaid on Figure~\ref{Vela_Jr} are H.E.S.S. \tev
contours of Vela Jr.  \citep{vela_jr_hess} and show that three
point-like catalog sources 2FGL\,J0853.5$-$4711, 2FGL\,J0848.5$-$4535,
and 2FGL\,J0855.4$-$4625 correlate with the multiwavelenth emission of
this SNR are were most likely included in 2FGL to account for residual
induced by not taking the extension of this extended source into account.
These sources were removed from our model of the sky.  In addition,
two further away sources 2FGL\,J0858.0$-$4815 and 2FGL\,J0901.7$-$4655
were removed from our model of the sky because they were not significant
above 10 \gev.

With this model of the sky, 2FGL\,J0851.7$-$4635 was found to have an
extension of $1\fdg13\pm0\fdg08_\stat\pm0\fdg05_\sys$ with an extension
significance of $\tsext=87.2$ using photons with energies between 10
\gev and 100 \gev.  This source was fit to $(l,b)=(266\fdg29,-1\fdg43)$.
While fitting the extension of 2FGL\,J0851.7$-$4635, we iteartively
relocalized the position of the nearby point-like catalog source
2FGL\,J0854.7$-$4501 to $(l,b)=(266\fdg24,0\fdg49)$ to better fit its
position at high energies.  The LAT size matches well the \tev morphology
of Vela Jr.  A detailed papers by the LAT Collaboration on this SNR is
in preparation.
}

@@

\tablenotetext{(a)}{
Integral Flux is in units of $10^{-9}\phflux$ and integrated in the fit
energy range (either 1 \gev 100 \gev or 10 \gev 100 \gev).
}

@@

\clearpage
\begin{figure}
    \ifcolorfigure
      \plotone{summary_plots/gev_vs_tev_plot_color.eps}
    \else
      \plotone{summary_plots/gev_vs_tev_plot_bw.eps}
      \fi
    % this plot came from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/compare_tev_gev_sizes/v3/run.py
    \caption{
    A comparison of the size of extended sources detected at
    both \gev and \tev energies.  The \tev size of W30,
    2FGL\,J1837.3$-$0700c, 2FGL\,J1632.4$-$4753c, 2FGL\,J1615.0$-$5051,
    and 2FGL\,J1615.2$-$5138 are from \cite{hess_plane_survey}.
    The \tev size of MSH\,15$-$52, HESS\,J1825$-$137,
    Vela X, Vela Jr., RX\,J1713.7$-$3946 and W28 are from
    \cite{msh_15_52_hess,hess_j1825_hess,vela_x_hess,vela_jr_hess,rx_j1713_hess,w28_hess}.
    The \tev size of IC443 is from \cite{ic443_veritas} and
    W51c is from \cite{w51c_with_magic_at_fermi_symposium}.  The \tev
    size of MSH\,15$-$52, HESS\,J1614$-$518, HESS\,J1632$-$478, and
    HESS\,J1837$-$069 have only been reported with an elliptical 2D
    Gaussian fit and so the assumed size is the average of the semi-major
    and semi-minor axis.
    The LAT extension of
    Vela X is from \cite{velax}.  Except for RX\,J1713.7$-$3946 and Vela Jr.,
    the \tev sources were fit assuming a 2D Gaussian surface brightness
    so the plotted \gev and \tev extensions were first converted to
    \rsixeight (see Section~\ref{compare_source_size}).  Because of
    their large size, the shape of RX\,J1713.7$-$3946 and Vela Jr.
    were not directly fit at \tev energies. Here, we take their sizes to be the
    same as the LAT size and do not converted them to \rsixeight. The LAT
    extension errors are the statistical and systematic errors added
    in quadrature. 
}\label{gev_vs_tev_plot}
  \end{figure}

\clearpage
\begin{figure}
    \ifcolorfigure
      \plotone{summary_plots/gev_vs_tev_histogram_color.eps}
    \else
      \plotone{summary_plots/gev_vs_tev_histogram_bw.eps}
    \fi
    % this plot came from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/extension_histogram/v4/extension_histogram.py
    \caption{
    The distribution of the size of the 22 extended LAT sources
    at \gev energies
    (colored blue in the electronic version) and the size of the
    42 extended H.E.S.S. sources at \tev energies
    (colored red).  The size of Vela X
    is taken from \cite{velax}.  Except for RX\,J1713.7$-$3946 and
    Vela Jr., the H.E.S.S. sources were fit with a 2D Gaussian surface
    brightness model so the LAT and H.E.S.S. sizes are first converted
    to \rsixeight. Because
    the spatial morphology of RX\,J1713.7$-$3946 and Vela Jr. is poorly
    described by a 2D Gaussian surface brightness model, the \gev and \tev
    extensions are included assuming a uniform surface brightness model.
    }\label{gev_vs_tev_histogram}
  \end{figure}

>>>>

@@

\newif\ifcolorfigure
%\colorfiguretrue
\colorfigurefalse

@@

\author{
J.~Lande\altaffilmark{2,1}, 
S.~Funk\altaffilmark{2,1}, 
M.~Ackermann\altaffilmark{3,1}, 
%A.~Allafort\altaffilmark{2},
%M.-H.~Grondin\altaffilmark{53},
% Keith Bechtol
% Marshall Roth?
% Marianne Lemoine-Goumard
% Francesco Giordano
% Toby Burnett
% Jean Ballet
% Matthew Kerr
\ldots
\altaffiltext{1}{Corresponding authors: M.~Ackermann, markus.ackermann@desy.de; S.~Funk, funk@slac.stanford.edu; J.~Lande, joshualande@gmail.com.}
\altaffiltext{2}{W. W. Hansen Experimental Physics Laboratory, Kavli Institute for Particle Astrophysics and Cosmology, Department of Physics and SLAC National Accelerator Laboratory, Stanford University, Stanford, CA 94305, USA}
\altaffiltext{3}{Deutsches Elektronen Synchrotron DESY, D-15738 Zeuthen, Germany}
}

\begin{abstract}
Spatial extension is an important characteristic for correctly
associating LAT sources with their counterparts at other wavelengths
and for obtaining an unbiased model of their spectra.  We present a new
method for quantifying the spatial extension of sources detected by the Large
Area Telescope (LAT), the primary science instrument on the {\em \fermi
Gamma-ray Space Telescope} (\fermi).  We perform a series of Monte Carlo
simulations to validate this tool and calculate the LAT threshold for
detecting the spatial extension of sources.  We then test all sources
in the second \fermi-LAT catalog (2FGL) for extension. 
\newtext{
We report the
detection of seven spatially extended sources in addition to the twelve
spatially extended sources reported in 2FGL and two others studied
in dedicated publications.
}
\end{abstract}

@@

\newtext{
Previous analysis of extended LAT sources was performed as dedicated
studies of individual sources so we expect that a systematic scan of
all LAT detected sources could uncover additional spatially extended
sources.  For these reasons, in Section~\ref{analysis_methods_section}
we present a new systematic method for analyzing spatially extended
LAT sources.  In Section~\ref{monte_carlo_validation}, we demonstrate
that this method can be used to test the statistical significance of the
extension of a LAT source and in Section~\ref{extension_sensitivity}
we calculate the LAT detection threshold to resolve the extension
of a source.  In Section~\ref{dual_localization_method}, we
study the ability of the LAT to distinguish between a spatially
extended source and the confusion of nearby point-like sources.
In Section~\ref{test_2lac_sources}, we further demonstrate that our
detection method does not misidentify point-like sources as being
extended by testing the extension of active galactic nuclei (AGN)
believed to be unresolvable.  In Section~\ref{validate_known},
we systematically reanalyze the twelve extended sources included
in 2FGL and in Section~\ref{systematic_errors_on_extension}
we describe a way to estimate systematic errors on the extension of a source.
In Section~\ref{extended_source_search_method}, we
describe a search for new spatially extended LAT sources. Finally,
in Section~\ref{new_ext_srcs_section} we present the detection of the
extension of nine spatially extended sources that were reported in 2FGL
but as being point-like.  Two of these sources have been previously analyzed in
dedicated publications.
}

@@

\newtext{
We considered two different models that can be used to model the
surface brightness profile for extended sources, either a 2D Gaussian
profile
\begin{equation}\label{gauss_pdf}
  I_\text{Gaussian}(x,y)=\tfrac{1}{2\pi\sigma^2}\exp\left(-(x^2+y^2)/2\sigma^2\right)
\end{equation}
or a uniform surface brightness profile
\begin{equation}\label{disk_pdf}
  I_\text{disk}(x,y)=
  \begin{cases}
    \frac{1}{\pi\sigma^2} & x^2+y^2\le\sigma^2 \\
    0                      & x^2+y^2>\sigma^2.
  \end{cases}
\end{equation}
Although these shapes are significantly different, 
Figure~\ref{compare_disk_gauss} shows that their PDF 
are similar for a source that has a 
0\fdg5
radius
typical of LAT detected extended sources.
To
allow a valid comparison between the Gaussian and the uniform surface
brightness morphologies,
we define the source size 
as the radius containing
68\% of the intensity ($\rsixeight$).  $\rsixeight_\text{,Gaussian}=1.51\sigma$
and $\rsixeight_\text{,disk}=0.82\sigma$ where $\sigma$
is defined in Equation~\ref{gauss_pdf} and
Equation~\ref{disk_pdf} respectively.
For this example, $\rsixeight=0\fdg5$ so
$\sigma_\text{disk}=0.61\degree$ and 
$\sigma_\text{Gaussian}=0.33\degree$.

For smaller and fainter extended sources, the differences in the PDF for
different spatial models lost in the noise and the LAT is not sensitive
to the detailed spatial structure of these sources.  Therefore, in our
search for extended sources we use only a radially-symmetric uniform
surface brightness as our spatial extension shape. Unless otherwise noted,
we quote the radius of the edge ($\sigma$) as the size of the source.
}

@@

Figure~\ref{index_sensitivity} shows the threshold for sources of four
spectral indices from 1.5 to 3 and extensions varying from $\sigma=0\fdg1$
to $2\fdg0$.  
The LAT flux threshold for a significant detection
of source extension drops quickly with
increasing source size and reaches a minimum around 0\fdg5. 
\newtext{
The threshold is high for small extensions when the
source is 
small compared to the size of the PSF. The threshold increases
for large extended sources because the source becomes
increasingly diluted by the background.
} Figure~\ref{index_sensitivity} shows
the threshold using photons with energies between 100 \mev and 100 \gev
and also using only photons with energies between 1 \gev and 100 \gev.
Except for very large or very soft
sources, our detection threshold is
not substantially improved by including photons with energies between 100 \mev and
1 \gev.  This is also demonstrated in Figure~\ref{four_plots_ic443}
which shows \tsext for the SNR IC443 computed independently in twelve
energy bins between 100 \mev and 100 \gev. For IC443, which has a
spectral index $\sim2.4$ and an extension $\sim0\fdg35$, 
almost the entire increase in likelihood
modeling the source as being extended comes
from energies above 1 \gev.  Furthermore, other systematic errors
become increasingly important at low energy. For our extension search
(Section~\ref{extended_source_search_method}),
we therefore used only photons with energies above 1 \gev.

@@

We assessed the power of the $\tsext>\tsinc$ test with a Monte Carlo
study.  We simulated one spatially extended source and fit it as both
an extended source and as two point-like sources.  We then simulated two
point-like sources and fit them with the same two hypothesis. By comparing
the distribution of \tsinc and \tsext for the two cases, we evaluated
how well the $\tsext>\tsinc$ test is at rejecting the possibility of
source confusion as well as how likely it is to incorrectly reject 
with this test that an extendd source is 
spatially extended.  All sources were simulated using
the same two-year time range included in 2FGL and all fit values were
calculated using \pointlike. The simulations were performed against
an background 10 times the isotropic background measured by EGRET,
representative of the background near the Galactic plane.

We did this study first in the energy range from 1 \gev to 100 \gev by
simulating extended sources of flux $4\times10^{-9} \phflux$ integrated
from 1 \gev to 100 \gev with a power-law spectral model with
spectral index 2.  This flux was picked to be representative of the
new extended sourced that were discovered in the following analysis
when looking in the 1 \gev to 100 \gev energy range
(see Section~\ref{new_ext_srcs_section}).
We simulated these sources using a uniform surface brightness spatial
model with extensions varying
up to $1\degree$.  Figure~\ref{confusion_plot}a shows 
the distribution of \tsinc and \tsext as a
function of the simulated extension of the source
for 200 statistically independent simulations.

Figure~\ref{confusion_plot}c shows the same analysis but when
fitting simulations of two point-like sources each with half of the
flux of the spatially extended source and with the same spectral index
as the extended source.
Finally, figure~\ref{confusion_plot}e shows the same analysis with each
point-like source having the same flux but different spectral indices.
One point-like source had a spectral index of 1.5 and the other an index
of 2.5.

@@

The distributions also shows that for simulated
extended sources with typical
sizes ($\sigma\sim0\fdg5$), one can often obtain almost as good an
increase in likelihood
when fitting the extended source as two point-like sources.
This is because although
two point-like sources represent an incorrect spatial model, 
the two point-like sources have four additional degrees
of freedom (two spatial and two spectral parameters)
than one extended source (one extension parameter)
and can therefore easy model
the extended source and statistical fluctuations in the data.
This effect is most pronounced when using photons with energies between 1
\gev and 100 \gev where the PSF is even larger.  
On the other hand,
when testing the two simulated point-like sources for extension our test
did a better job at separating the distribution of \tsinc and \tsext
and found that \tsinc was typically much larger than \tsext.
This is because the one additional extension parameter is less able to
model the second source or statistical fluctuations.

From this, we conclude that the $\tsext>\tsinc$ test is 
powerful at discarding cases in which the true emission is derived
from two point-like sources instead of from a spatially extended
source. On the other hand, because of the large values of \tsinc
in figure~\ref{confusion_plot}a and figure~\ref{confusion_plot}b when
fitting spatially extended sources, we caution that even in the case of
their being spatially extended emission the fit of two point-like sources
can often seem (by comparing the likelihoods) to be almost as good.

Finally, figure~\ref{index_difference} shows the absolute value of the
difference of the fit spectral indices for the case of fitting simulated
spatially extended sources as two point-like sources. The plot compares
the fit in the 1 \gev to 100 \gev energy range and in the 10 \gev to
100 \gev energy range. An error on the difference in spectral index is
calculated by adding in quadrature the statistical errors on the spectral
indices.  This plot shows that there can be a large difference in the
spectral index of the two fit point-like sources which is often greater
than the statistical errors on the difference in spectral index. This is
because the fit values the errors on an incorrect model are not expected
to be meaningful. Therefore, the difference in spectral index when fitting
an extended source candidate as two point-like sources can not be easily
used to help determine whether a source is actually spatially extended
instead of being the emission of two nearby point-like sources.

From this Monte Carlo study, we can see the limits of a LAT analysis of
spatially extended sources.  Section~\ref{monte_carlo_validation} showed
that we have a statistical test that finds when a LAT source is
not well described by the PSF. 
But this test does
not uniquely prove that the emission originates from spatially extended
emission instead of from multiple unresolved sources.  Demanding that
$\tsext>\tsinc$ is a powerful second test to avoid cases of simple
confusion of two point-like sources. But it could always be the
case that an extended source candidate is actually the superposition
of multiple point-like or extended sources that could be resolved
with deeper observations of the region.  There is nothing
about this conclusion unique to LAT analysis, but the
larger PSF of the LAT and the density of sources expected to be \gev emitters in
the galactic plane makes this issue more significant for LAT analysis.
When possible, multiwavelength information should be used to help 
select the best model of the sky.

@@

\newtext{
2FGL\,J0851.7$-$4635 was found by our search to be an extended source
candidate in the 10 \gev to 100 \gev energy range and is spatially
coincident with the SNR Vela Jr and was recently studied by the LAT
collaboration in \cite{vel_jr_lat}.  Figure~\ref{Vela_Jr} shows a counts
map of the source.  Overlaid on Figure~\ref{Vela_Jr} are H.E.S.S. \tev
contours of Vela Jr.  \citep{vela_jr_hess} and show that three
point-like catalog sources 2FGL\,J0853.5$-$4711, 2FGL\,J0848.5$-$4535,
and 2FGL\,J0855.4$-$4625 correlate with the multiwavelenth emission of
this SNR are were most likely included in 2FGL to account for residual
induced by not taking the extension Vela Jr. into account.  We removed
from our model of the sky these sources.  In addition, two further away
sources 2FGL\,J0858.0$-$4815 and 2FGL\,J0901.7$-$4655 were removed from
our model of the sky because they were not significant above 10 \gev.

With this model of the sky, 2FGL\,J0851.7$-$4635 was found to have an
extension of $1\fdg13\pm0\fdg08_\stat\pm0\fdg05_\sys$ with an extension
significance of $\tsext=87.2$.
The LAT size matches well the \tev morphology
of Vela Jr. This source was fit to $(l,b)=(266\fdg29,-1\fdg43)$.
While fitting the extension of 2FGL\,J0851.7$-$4635, we iteratively 
relocalized the position of the nearby point-like catalog source
2FGL\,J0854.7$-$4501 to $(l,b)=(266\fdg24,0\fdg49)$ to better fit its
position at high energies.  
}

@@

\tablenotetext{(a)}{
Integral Flux in units of $10^{-9}\phflux$ and integrated in the fit
energy range (either 1 \gev 100 \gev or 10 \gev 100 \gev).
}

@@

\clearpage
\begin{figure}
    \ifcolorfigure
      \plotone{summary_plots/gev_vs_tev_plot_color.eps}
    \else
      \plotone{summary_plots/gev_vs_tev_plot_bw.eps}
      \fi
    % this plot came from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/compare_tev_gev_sizes/v4/run.py
    \caption{
    A comparison of the size of extended sources detected at
    both \gev and \tev energies.  The \tev size of W30,
    2FGL\,J1837.3$-$0700c, 2FGL\,J1632.4$-$4753c, 2FGL\,J1615.0$-$5051,
    and 2FGL\,J1615.2$-$5138 are from \cite{hess_plane_survey}.
    The \tev size of MSH\,15$-$52, HESS\,J1825$-$137,
    Vela X, Vela Jr., RX\,J1713.7$-$3946 and W28 are from
    \cite{msh_15_52_hess,hess_j1825_hess,vela_x_hess,vela_jr_hess,rx_j1713_hess,w28_hess}.
    The \tev size of IC443 is from \cite{ic443_veritas} and
    W51c is from \cite{w51c_with_magic_at_fermi_symposium}.  The \tev
    size of MSH\,15$-$52, HESS\,J1614$-$518, HESS\,J1632$-$478, and
    HESS\,J1837$-$069 have only been reported with an elliptical 2D
    Gaussian fit and so the assumed size is the average of the semi-major
    and semi-minor axis.
    The LAT extension of
    Vela X is from \cite{velax}. 
    The \tev sources were fit assuming a 2D Gaussian surface brightness
    so the plotted \gev and \tev extensions were first converted to
    \rsixeight (see Section~\ref{compare_source_size}).  
    \newtext{Because of
    their large size, the shape of RX\,J1713.7$-$3946 and Vela Jr.
    were not directly fit at \tev energies and so are not included
    in this comparison. On the other hand, dedicated LAT publications on
    these sources showed that their morphologies are consistent
    \citep{rx_j1713_lat,vela_jr_lat}.
    }
    The LAT
    extension errors are the statistical and systematic errors added
    in quadrature. 
}\label{gev_vs_tev_plot}
  \end{figure}

\clearpage
\begin{figure}
    \ifcolorfigure
      \plotone{summary_plots/gev_vs_tev_histogram_color.eps}
    \else
      \plotone{summary_plots/gev_vs_tev_histogram_bw.eps}
    \fi
    % this plot came from /u/gl/lande/work/fermi/extended_catalog/2FGL/plots_for_paper/extension_histogram/v4/extension_histogram.py
    \caption{
    The distribution of the size of the 22 extended LAT sources
    at \gev energies
    (colored blue in the electronic version) and the size of the
    42 extended H.E.S.S. sources at \tev energies
    (colored red).  The size of Vela X
    is taken from \cite{velax}.  Except for RX\,J1713.7$-$3946 and
    Vela Jr., the H.E.S.S. sources were fit with a 2D Gaussian surface
    brightness model so the LAT and H.E.S.S. sizes are first converted
    to \rsixeight. 
    \newtext{
    Because of their large size, the shape of RX\,J1713.7$-$3946 and
    Vela Jr. were not directly fit at \tev energies. Their size at
    \tev energies is taken from the HESS source catalog to be 0\fdg5
    and $1\degree$ respectively.  Because the spatial morphology of
    RX\,J1713.7$-$3946 and Vela Jr. is poorly described by a 2D Gaussian,
    the \gev and \tev extensions of these sources was not converted
    to \rsixeight.
    }
    }\label{gev_vs_tev_histogram}
  \end{figure}
